Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} Model System

### 2014-05-30 paper_enahpc14/Sec1:Related.tex
cleans up latest version; improves format
updates a few things according to the reviewers' suggestions

\emph{Ou  Pang et  al.}   \cite{Ou-2012} compare  ARM  and Intel  x86
workstations  clusters  and   conclude  that  ARM-based  clusters  are
advantageous  with   lightweight  applications  in   terms  of  energy
efficiency.

\emph{G\"oddeke et al.}  \cite{Goddeke-2013} evaluate weak and strong
scalability of PDE solvers on a cluster of 96 ARM dual-core processors
and demonstrate  that the ARM-based  cluster can be more  efficient in
terms of energy-to-solution compared to x86-based cluster.

@@

% \emph{Cumming  et  al.}   \cite{Cumming-2014}  present a  simple  and
% practical  methodology  looking at  energy  minimization  that can  be
% applied to various applications.

>>>>

% \emph{Ou  Pang et  al.}   \cite{Ou-2012} compare  ARM  and Intel  x86
% workstations  clusters  and   conclude  that  ARM-based  clusters  are
% advantageous  with   lightweight  applications  in   terms  of  energy
% efficiency.

% \emph{G\"oddeke et al.}  \cite{Goddeke-2013} evaluate weak and strong
% scalability of PDE solvers on a cluster of 96 ARM dual-core processors
% and demonstrate  that the ARM-based  cluster can be more  efficient in
% terms of energy-to-solution compared to x86-based cluster.

@@

\emph{Cumming  et  al.}  \cite{Cumming-2014}  present  a simple  and
practical  methodology  looking at  energy  minimization  that can  be
applied to various applications.

\emph{Diouri et al.} \cite{suscom} present a deep 
study of different internal and external power monitoring devices that
are widely used nowadays.

### 2014-05-30 paper_enahpc14/Sec3:Environment-setup.tex

Benchmarks presented  in this work  were performed on  three different
platforms  and can  be reproduced  within an  expected  variance, thus
allowing a  fair comparison between the baseline  and future milestone
versions of \cosmoart.

@@

\item \tinto is  a heterogeneous cluster composed of  28 compute nodes
  at UJI.  For our experiments only  a subset of  16 homogeneous nodes
  was  considered. Each  node  is  comprised of  two  Intel Xeon Westmere EP E5645
  hexa-core processors  running at  2.4\,GHz, equipped with  24\,GB of
  DDR3  1333\,MHz and  connected via  InfiniBand QDR  with  a Mellanox
  MTS\-3600 switch. Power saving mechanisms are enabled on the nodes are the C-states C1, C3 and C6, 
  but no P-states (except P0) are used. TurboBoost feature is enabled for our experiments.
\end{itemize}

@@

Supercomputer clusters  considered for our experiments at  CSCS of ETH
Zurich are  equipped with E3METER  Intelligent Power Strips  (IPS) and
Monitors (IPM)  which are high quality electricity  meters released by
Riedo Networks,  that enable to  monitor and log power  consumption of
the  IT infrastructure  as well  as constantly  analyze  line voltage,
current, power-factor, frequency  with $\pm1\,\%$ accuracy.  Using reliable
narrowband power-line communication (PLC) technology, all metering and
power  quality data  from  each  IPS are  centrally  collected by  the
E3METER Data Concentrator, via the existing power cables thus avoiding
the need  for extra  cabling.  This data  is made available  via SNMP,
HTTP  and  TELNET  through  the  built-in Fast  Ethernet  port.   Time
synchronization is guaranteed by  using NTP servers.  Measured data is
accessed through the open  source Cacti software including the E3METER
Cacti Plugin to  scan the entire PLC network  and monitor in real-time
the power  usage of  individual rack, recorded  in 5  minutes interval
periods.

@@

To employ our  framework, \cosmoart is first compiled  and linked with
the  Extrae instrumentation  library, which  automatically instruments
the Fortran~90  code of the model,  e.g., to record  MPI calls.  Next,
\cosmoart  is run  on the  nodes, thus  consuming certain  amount of
power. The  nodes are connected  to APC 8653 Power  Distribution Units
(PDUs) that account  for the consumed   power with  a sampling rate of
1\,Hz and $\pm3\,\%$ accuracy. The power data captured by  these PDUs is collected by a \pmlib
tracing server running  on a separate machine. The  client, running on
the target nodes, sends start/stop  primitives in order to collect the
power  data during  the \cosmoart  run. Once  the run  is  finished, a
power-performance  trace file  containing the  power data  is received
from  the  tracing  server  and the  instrumentation  post-process  is
generated.  A  merged trace  file  is  finally  loaded in  Paraver  to
visualize  and correlate  the  performance trace  and  with the  power
profile  generated by \cosmoart.  This procedure  allows us  to easily
detect power bottlenecks, basically constituted by inefficient waiting
methods.

>>>>

Benchmarks presented  in this work  were performed on  three different
platforms  and can  be reproduced  within an  expected  variance, thus
allowing a  fair comparison between the baseline  and future milestone
versions of \cosmoart:

@@

\item \tinto is  a heterogeneous cluster composed of  28 compute nodes
  at UJI.  For our experiments only  a subset of  16 homogeneous nodes
  was  considered. Each  node  is  comprised of  two  Intel Xeon Westmere EP E5645
  hexa-core processors  running at  2.4\,GHz, equipped with  24\,GB of
  DDR3  1333\,MHz and  connected via  InfiniBand QDR  with  a Mellanox
  MTS\-3600 switch. 
\end{itemize}

Power saving mechanisms enabled on all the nodes' platforms are the C-states (C1, C3 and C6) 
and P-states, but only P0 or the maximum frequency is used. TurboBoost feature is also enabled for our experiments.

@@

Supercomputer clusters  considered for our experiments at  CSCS of ETH
Zurich are  equipped with E3METER  Intelligent Power Strips  (IPS) and
Monitors (IPM)  which are high quality electricity  meters released by
Riedo Networks,  that enable to  monitor and log power  consumption of
the  IT infrastructure  as well  as constantly  analyze  line voltage,
current, power-factor, frequency  with $\pm1\,\%$ accuracy.  Using reliable
narrowband power-line communication (PLC) technology, all metering and
power  quality data  from  each  IPS are  centrally  collected by  the
E3METER Data Concentrator, via the existing power cables thus avoiding
the need  for extra  cabling.  This data  is made available  via SNMP,
HTTP  and  TELNET  through  the  built-in Fast  Ethernet  port.   Time
synchronization is guaranteed by  using NTP servers. Averaged measured data is
accessed through the open  source Cacti software including the E3METER
Cacti Plugin to  scan the entire PLC network  and monitor in real-time
the power  usage of  individual rack, recorded  in 5  minutes interval
periods.

@@

To employ our  framework, \cosmoart is first compiled  and linked with
the  Extrae instrumentation  library, which  automatically instruments
the Fortran~90  code of the model,  e.g., to record  MPI calls.  Next,
\cosmoart  is run  on the  nodes, thus  consuming certain  amount of
power. The  nodes are connected  to APC 8653 Power  Distribution Units
(PDUs) that account  for the consumed   power with  a sampling rate of
1\,Hz and $\pm3\,\%$ accuracy. The power data captured by  these PDUs is collected by a \pmlib
tracing server running  on a separate machine. The  client, running on
the target nodes, sends start/stop  primitives in order to collect the
power  data during  the \cosmoart  run. Once  the run  is  finished, a
power-performance  trace file  containing the  power data  is received
from  the  tracing  server  and the  instrumentation  post-process  is
generated.  A  merged trace  file  is  finally  loaded in  Paraver  to
visualize  and correlate  the  tasks trace with the  power
profile  produced by \cosmoart.  This procedure  allows us  to easily
detect power bottlenecks, basically constituted by inefficient waiting
methods.

### 2014-05-30 paper_enahpc14/paper.tex

\begin{abstract}
  In this  paper we investigate  the energy footprint  and performance
  profiling of \cosmoart  on various HPC platforms.  This  model is an
  extension of  the operational weather  forecast model of  the German
  weather  service   (DWD),  developed  for  the   evaluation  of  the
  interactions of reactive gases  and aerosol particles with the state
  of atmosphere at the  regional scale.  Different measurement devices
  and energy-aware techniques are  described to evaluate both time and
  energy  to  solution  of  the  considered application  and  to  gain
  detailed  insights  into power  and  performance requirements.   Our
  motivation  is to  improve  corresponding code  sections to  sustain
  performance  while minimising energy-to-solution.   This preliminary
  work  sets the  basis for  subsequent studies  to  tackle challenges
  related  to  energy  efficient  high performance  computing  in  the
  framework of the Exa2Green project \cite{EXA2GREEN}.

\keywords{High performance computing  \and Energy-aware computing
  %Green computing  
  \and Numerical weather  prediction \and Atmospheric
  chemistry  \and  Aerosol  modelling \and
  Benchmark  analysis  \and 
  \cosmoart  coupled  model \and  
  Power-performance profiling and tracing tools  \and
  Energy-saving techniques}
% \PACS{PACS code1 \and PACS code2 \and more} 
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

\begin{acknowledgements}
The research  leading to these  results is supported by  the Exa2Green
project  co-financed by  the European  Commission under  7th Framework
Programme Future and Emerging Technologies (FET) Proactive Initiative:
Minimising Energy  Consumption of Computing to the  Limit (MINECC). We
also gratefully acknowledge the High Performance and High Productivity
Computing Initiative \cite{HP2C} for results that will be leveraged in
subsequent code refactoring.
\end{acknowledgements}

>>>>

\begin{abstract}
  In this  paper we investigate  the energy footprint  and performance
  profiling of \cosmoart  on various HPC platforms.  This  model is an
  extension of  the operational weather forecast model of  the German
  wea\-ther  service   (DWD),  developed  for  the   evaluation  of  the
  interactions of reactive gases  and aerosol particles with the state
  of atmosphere at the  regional scale.  Different measurement devices
  and energy-aware techniques are  described to evaluate both time and
  energy  to  solution  of  the  considered application  and  to  gain
  detailed  insights  into power  and  performance requirements.   Our
  motivation  is to  improve  corresponding code  sections to  sustain
  performance  while minimizing energy-to-solution.   This preliminary
  work  sets the  basis for  subsequent studies  to  tackle challenges
  related  to  energy  efficient  high performance  computing  in  the
  framework of the Exa2Green project \cite{EXA2GREEN}.

\keywords{High performance computing  \and Energy-aware computing
  %Green computing  
  \and Numerical weather  prediction \and Atmospheric
  chemistry  \and  Aerosol  modeling \and
  Benchmark  analysis  \and 
  \cosmoart  coupled  model \and  
  Power-performance profiling and tracing tools  \and
  Energy-saving techniques}
% \PACS{PACS code1 \and PACS code2 \and more} 
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

\begin{acknowledgements}
The research  leading to these  results is supported by  the Exa2Green
project  co-financed by  the European  Commission under  7th Framework
Programme Future and Emerging Technologies (FET) Proactive Initiative:
Minimizing Energy  Consumption of Computing to the  Limit (MINECC). We
also gratefully acknowledge the High Performance and High Productivity
Computing Initiative \cite{HP2C} for results that will be leveraged in
subsequent code refactoring.
\end{acknowledgements}
